Two Decades of Smalltalk VM Development

### 2018-08-17 paper.tex
Feenk=>feenk
I missed one: the simulator slow performance => slow simulator performance.
minor disambiguation.

%% Author with single affiliation.
\author{Eliot Miranda}
                                        %% can be repeated if necessary
\affiliation{
  %\position{Position1}
 % \department{VM team}              %% \department is recommended
  \institution{Feenk}            %% \institution is required
 % \streetaddress{Street1 Address1}
  \city{San Francisco}
  %\state{France}
  %\postcode{Post-Code1}
  \country{California}                    %% \country is recommended
}
\email{eliot.miranda@gmail.com}          %% \email is recommended

@@

This code is the end of the entry sequence that checks the stack pointer against the stack limit for the end of the current stack page, which is also used to break out of machine code to service interrupts or events such as invoking the scavenger, etc.  \emph{\%rbx} is used as a base register pointing at the variables in interpreter.  Following that is the code for \emph{index := self indexOf: anElement startingAt: start} where \emph{anElement} and \emph{start} are method parameters and \emph{index} is a local variable.  HasBytecodePC and IsSendCall are decorations that identify important locations in machine code, including object references, runtime calls, etc. These points are specified using metadata implemented as a simple bytecoded language, a stream of such bytes being added to the end of each machine code method.  This metadata is parsed when the garbage collector needs to locate object references in machine code, when linked sends must be relocated when the machine code zone is compacted to make space by throwing away some number of least recently used methods, and when machine code pcs must be mapped to their corresponding bytecode pcs since at the Smalltalk level only bytecode pcs are seen. %to preserve the illusion that the VM is a bytecode interpreter using a spaghetti stack of Context objects.

@@

Although the RPython code can be executed as normal Python code, the developers seem to think it is not worth to do such a thing, mostly because executing code in this way is very slow. 
The overall architecture of the RPython toolchain is different, resulting in a longer time to reach peak performance (though their peak performance is at least in theory better than with our JIT). This time may be quite significant in simulation mode. In contrast, in \openSmalltalkVM, we can reuse the snapshots to work around the simulation slow performance.

>>>>

%% Author with single affiliation.
\author{Eliot Miranda}
                                        %% can be repeated if necessary
\affiliation{
  %\position{Position1}
 % \department{VM team}              %% \department is recommended
  \institution{feenk}            %% \institution is required
 % \streetaddress{Street1 Address1}
  \city{San Francisco}
  %\state{France}
  %\postcode{Post-Code1}
  \country{California}                    %% \country is recommended
}
\email{eliot.miranda@gmail.com}          %% \email is recommended

@@

This code is the end of the entry sequence that checks the stack pointer against the stack limit for the end of the current stack page, a check which is also used to break out of machine code to service interrupts or events such as invoking the scavenger, etc.  \emph{\%rbx} is used as a base register pointing at the variables in interpreter.  Following that is the code for \emph{index := self indexOf: anElement startingAt: start} where \emph{anElement} and \emph{start} are method parameters and \emph{index} is a local variable.  HasBytecodePC and IsSendCall are decorations that identify important locations in machine code, including object references, runtime calls, etc. These points are specified using metadata implemented as a simple bytecoded language, a stream of such bytes being added to the end of each machine code method.  This metadata is parsed when the garbage collector needs to locate object references in machine code, when linked sends must be relocated when the machine code zone is compacted to make space by throwing away some number of least recently used methods, and when machine code pcs must be mapped to their corresponding bytecode pcs since at the Smalltalk level only bytecode pcs are seen. %to preserve the illusion that the VM is a bytecode interpreter using a spaghetti stack of Context objects.

@@

Although the RPython code can be executed as normal Python code, the developers seem to think it is not worth to do such a thing, mostly because executing code in this way is very slow. 
The overall architecture of the RPython toolchain is different, resulting in a longer time to reach peak performance (though their peak performance is at least in theory better than with our JIT). This time may be quite significant in simulation mode. In contrast, in \openSmalltalkVM, we can reuse the snapshots to work around slow simulation performance.
